(sec:how-to-access)=
# Access from Python, R, MATLAB, etc.


The current version of the benchmark dataset battery can be downloaded
from <https://github.com/gagolews/clustering-data-v1/releases/tag/v1.1.0>.

## Python

To facilitate the comparison of clustering algorithms in the
Python environment, we have developed a dedicated package
that is available for download from
[PyPI](https://pypi.org/project/clustering-benchmarks/);
see {ref}`sec:clustbench-usage` for more details.

However, as described in the {ref}`sec:file-format` section,
all data files are neat and tidy. Therefore, we can access them
easily using some more low-level functions from *numpy*. For example:

```{python}
import numpy as np
import os.path
# to do: change to your local path
base_name = os.path.join("~", "Projects", "clustering-data-v1", "wut", "smile")
base_name = os.path.expanduser(base_name)
data    = np.loadtxt(base_name + ".data.gz", ndmin=2)
data[:6, :]  # preview
```

```{python}
labels  = np.loadtxt(base_name + ".labels0.gz", dtype="int")
labels[:6]  # preview
```


Note that cluster validity measures discussed in the
{ref}`Appendix <sec:external-validity-measures>`
are implemented in the [*genieclust*](https://genieclust.gagolewski.com)
package.


::::{note}
*To learn more about Python,
check out Marek's open-access (free!) textbook*
[Minimalist Data Wrangling in Python](https://datawranglingpy.gagolewski.com/)
{cite}`datawranglingpy`.
::::



## R

Following the {ref}`sec:file-format`,
the datasets can be accessed using the built-in R functions:

```{r}
# to do: change to your local path
base_name <- file.path("~", "Projects", "clustering-data-v1", "wut", "smile")
data    <- as.matrix(read.table(paste0(base_name, ".data.gz")))
head(data)  # preview
```

```{r}
labels  <- scan(paste0(base_name, ".labels0.gz"), integer())
head(labels)  # preview
```



{ref}`Cluster validity measures <sec:external-validity-measures>`
are implemented in the R version of the
[*genieclust*](https://genieclust.gagolewski.com) package.


R code can be called in Python using, for example,
the [*rpy2*](https://pypi.org/project/rpy2/) package.


::::{note}
*To learn more about R, check out Marek's open-access (free!) textbook*
[Deep R Programming](https://deepr.gagolewski.com/)
{cite}`deepr`.
::::




## MATLAB

Unfortunately, MATLAB is not free software.

It does not seem to be able to un*gzip* files
on the fly, but they can be decompressed to a temporary folder
manually.

```matlab
base_name = "~/Projects/clustering-data-v1/wut/smile";
t = tempdir();
data = readmatrix(char(gunzip(base_name + ".data.gz", t)), FileType="text");
labels = readmatrix(char(gunzip(base_name + ".labels0.gz", t)), FileType="text");
```

Note that there is also a MATLAB
[interface](https://au.mathworks.com/products/matlab/matlab-and-python.html)
for Python. This way, algorithms that have only been implemented in the
former can be called from within the latter.


<!--
## Julia

TODO: describe (help needed)
-->


::::{todo}
Contributions are welcome: Describe how to load
the datasets and benchmark results
in GNU Octave, Scilab, Julia, Mathematica, ... (🚧 help needed 🚧)
::::
